Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-08-08 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
updated paper.Rmd
added doc to binning functions

# Chr {{hot_chr}}

```{r, def-{{hot_chr}}}
hot_lower <- {{hot_mid}} - 2
hot_upper <- {{hot_mid}} + 2
fp <- paste0("Chr", {{hot_chr}}, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

@@

```{r, scols-{{hot_chr}}}
scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
```

@@

```{r, scols_extra-{{hot_chr}}}
scols_extra <- purrr::map(.x = local_annot2$symbol,
                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
) %>%
  bind_cols() %>%
  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
  mutate(local = scols1)
```

```{r, n-med-{{hot_chr}}}
n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
# only those local traits with n_med_tr or more mediated nonlocal traits 
# appear as annotation columns in the heatmap
n_med_per_local <- med_tib %>%
  group_by(local_symbol) %>%
  filter(lod_diff > 1.5) %>%
  count() %>%
  filter(n >= n_med_tr)
n_med_per_local %>%
  arrange(desc(n)) %>%
  knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
```

@@

(p <- pm[ord, ord] %>% # order both cols and rows by 'ord'
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
```

>>>>

## Chr {{hot_chr}}

```{r, def-{{hot_chr}}}
hot_lower <- {{hot_mid}} - 2
hot_upper <- {{hot_mid}} + 2
```

@@

```{r, scols-{{hot_chr}}}
scols1 <- tibble(local = 1L + (rownames(pm) %in% local_annot2$symbol), 
                 symbol = rownames(pm), 
                 mediated_by_keller_mediator = rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
                 )

```

@@

```{r, scols_extra-{{hot_chr}}}
scols_extra <- local_annot2$symbol %>%
  as.list() %>%
  (function(x) {names(x) <- unlist(x); x}) %>%
  purrr::map(.f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
) %>%
  bind_cols() %>%
  mutate(symbol = rownames(pm)) %>%
left_join(scols1, by = "symbol")
```

```{r, scols-peak-position-{{hot_chr}}}
annot_all <- local_annot2 %>%
  bind_rows(nonlocal_annot)

binning <- annot_all %>%
  dplyr::select(symbol, pos, lod) %>%
  mutate(bin = unlist(pos) %>%
           map(.f = function(x) {x %>% 
               bin_peak_position(breaks = hot_lower + (1:9) * 0.4) %>%
               binarize_bin_number(total_number_of_bins = 10L)
             })
  )
```

```{r, scan1-{{hot_chr}}}
fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1_out.rds")
if (file.exists(fn)) {
  s1out <- readRDS(fn)
} else {
  s1out <- qtl2::scan1(genoprobs = probs, 
                       pheno = cbind(hot_local, hot_nonlocal), 
                       kinship = K[["{{hot_chr}}"]], 
                       addcovar = covar, 
                       reml = TRUE
                       )
  saveRDS(s1out, fn)
}
```

```{r allele-effects-{{hot_chr}}}
fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1coef_out.rds")
if (file.exists(fn)) {
  s1coef_out <- readRDS(fn)
} else {
s1coef_out <- purrr::map(.x = tibble(cbind(hot_local, hot_nonlocal)), 
           .f = function(pheno) {
             qtl2::scan1coef(genoprobs = probs[["{{hot_chr}}"]], 
                             pheno = pheno, 
                             kinship = K[["{{hot_chr}}"]], 
                             addcovar = covar, 
                             reml = TRUE, 
                             )
           }
                )
saveRDS(s1coef_out, fn)
}
```

```{r, get-allele-effects-{{hot_chr}}}
# add index values for peak position
local_annot2 %>%
  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id)) %>%
  mutate(peak_effects = qtl2pleio::get_effects(peak_index, allele_effects_matrix = s1coef_out, map = map[["{{hot_chr}}"]]))

nonlocal_annot %>%
  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id))

```

```{r, n-med-{{hot_chr}}}
n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
# only those local traits with n_med_tr or more mediated nonlocal traits 
# appear as annotation columns in the heatmap
n_med_per_local <- med_tib %>%
  group_by(local_symbol) %>%
  filter(lod_diff > 1.5) %>%
  count() %>%
  filter(n >= n_med_tr)
n_med_per_local %>%
  arrange(desc(n)) %>%
  knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")
```

```{r define-scols-pre-{{hot_chr}}}
scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
```

@@

(p <- pm[ord, ord] %>% # order both cols and rows by 'ord'
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply::heatmapr(Rowv = FALSE,
            Colv = FALSE,
            plot_method = "plotly"
            ) %>%
    heatmaply::heatmaply_heatmapr2(
            fontsize_col = 2,
            fontsize_row = 2,
            symm = TRUE,
            row_side_colors = list(scols[ord, ], tibble(scols1$local[ord])),
            row_side_palette = list(heatmaply::cool_warm, heatmaply::Blues), 
            col_side_colors = tibble(local = scols1$local[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
```

### 2019-08-08 analysis/paper/Rmd/hotspots-paper-2.Rmd
added doc to binning functions

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "August 3, 2019"
bibliography: research.bib
linkReferences: true
link-citations: true 
output: 
  bookdown::html_document2: 
    code_folding: hide
    self_contained: no
    keep_md: true
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
---

@@

```{r readRDS}
lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
```

@@

`r if (eval_code) knitr::knit(text = out[[2]])`

`r if (eval_code) knitr::knit(text = out[[3]])`

`r if (eval_code) knitr::knit(text = out[[4]])`

`r if (eval_code) knitr::knit(text = out[[5]])`

>>>>

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "August 3, 2019"
bibliography: research.bib
linkReferences: true
link-citations: true 
output: 
  bookdown::html_document2: 
    code_folding: hide
---

@@

```{r readRDS}
lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
probs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
```

@@
